### 2018-02-07 appendix.tex
changed some notations in the appendix and did some edits for the regret proof (not fixed)

\section*{Proof of Lemma~\ref{lem:n0-inequality}}

\begin{proof}
First, we observe that
\begin{align}
&\frac{nx}{4\sigma}\geq \sqrt{0.6n\log(\log_{1.1}(n)+1)} \nonumber \\
\iff &\frac{n}{\log(\log_{1.1}(n)+1)}\geq \frac{9.6\sigma^2}{x^2}. \nonumber 
\end{align}
Since $\log(x)\leq x-1$ for $x>0$, we know 
\begin{align}
\log(\log_{1.1}(n)+1)=\log\left(\frac{\log(n)}{\log(1.1)}+1\right)\leq \log(11\log(n)+1)\leq \log(11n)\leq 3+\log(n). \nonumber
\end{align}

Thus, we know
\begin{align}
\frac{n}{\log(\log_{1.1}(n)+1)}\geq \frac{n}{3+\log(n)}. \nonumber 
\end{align}

To prove the lemma, we just need to show for $n\geq s_{0}$, we have
\begin{align}
\frac{n}{3+\log(n)}\geq \frac{9.6\sigma^2}{x^2}. \label{n0-equ}
\end{align}
Inequality~(\ref{n0-equ}) is true because of the following two observations:
\begin{itemize}
\item for $n\geq 50$, we have $\frac{n}{3+\log(n)}\geq n^{0.5}$;
\item for $n\geq \frac{92.16\sigma^4}{x^4}$, we have $n^{0.5}\geq {9.6\sigma^{2}}{x^{2}}$.
\end{itemize}

Thus, we know our lemma is true.

\end{proof}

\section*{Proof of Theorem~\ref{rst:budget}}

We need the following lemma in part of the proof of Theorem~\ref{rst:budget}.
\begin{lemma}
For all $n\geq 1$, we have
\begin{align}
5n^{3/5} \geq \sqrt{0.6n \log(\log_{1.1}(n)+1)}. \nonumber
\end{align}
\label{lemma:cal2}
\end{lemma}
\begin{proof}
\begin{align}
&5 n^{3/5} \geq \sqrt{0.6n \log(\log_{1.1}(n)+1)} \nonumber \\
\Longleftarrow &  25 n^{6/5} \geq n \log(\log_{1.1}(n)+1) \nonumber  \\
\Longleftarrow &  25 n^{1/5} \geq \log(\log_{1.1}(n)+1) \nonumber \\
\Longleftarrow &  25 n^{1/5} \geq \log_{1.1}(n)+1. \nonumber 
\end{align}

Denote $f(x) = 25x^{1/5} - \log_{1.1}(x)-1$. Set it's first derivate to $0$, we get its global mimumum at $x_0=\frac{161051}{3125}$ and $f(x_0)> 0$. Thus, our Lemma holds true.

\end{proof}

Now we are ready to prove our first main result, Theorem~\ref{rst:budget}.

\begin{proof}

\noindent\textbf{Step 1: Categorize measurement errors into different radius envelopes}

Denote $\epsilon_{i,t}=\ArmEV{t}{i}-\ArmV{i}$ to be the estimation error for the attribute vector $u_i$ at time $t$. Denote $\epsilon_{i,t}^{j}$ to be the $j^{th}$ component of $\epsilon_{i,t}$. Denote $\omega$ to be a sample path and $n(t,\omega)$ to be the round number for sample path $\omega$ at time $t$. For a fixed time $t$, define
\begin{align}
L^{'}[\ell](t) = \{\omega:|\epsilon_{i,t}^{j}(\omega)|\leq g(n(t,\omega),\ell), \forall i,j\}\nonumber
\end{align}
where $g(n,\ell)$ is a function which we will define later. Define $L[1](t) = L^{'}[1](t)$ and $L[i](t) = L^{'}[i](t)\setminus L^{'}[i-1](t)$ for $i\geq 2$. We call $L[\ell](t)$ the $\ell^{th}$ envelope at time $t$. We often simplify the notation and use $L[\ell]$ instead of $L[\ell](t)$ without confusion.

In the calculation below, we omit the dependency on $\omega$ when refering to variables $c(t)$, $\epsilon_{i,t}^{j}$ and $t_n$. Based on the definition of $L[\ell]$, we know if $\omega\in L[\ell]$, the maximum payment we need to offer at time $t$ is bounded above by 
\begin{align}
&\max_{i}\AgentV{t}\cdot \ArmEV{t}{i} - \min_{j}\AgentV{t}\cdot \ArmEV{t}{j} \nonumber \\
= &\max_{i}\AgentV{t}\cdot (\epsilon_{i,t}+\ArmV{i}) - \min_{j}\AgentV{t}\cdot (\epsilon_{j,t}+\ArmV{j}) \nonumber \\
\leq &\max_{i}\AgentV{t}\cdot \ArmV{i} - \min_{j}\AgentV{t}\cdot \ArmV{j} +\max_{i}\AgentV{t}\cdot \epsilon_{i,t} - \min_{j}\AgentV{t}\cdot \epsilon_{j,t}\nonumber \\
\leq & R + 2\Diam dg(n,\ell). \nonumber
\end{align}

We denote $\bar{c}(n,\ell)=R + 2\Diam dg(n,\ell)$ as the upper bound of payment for round $n$ if our measurement error lie in the $\ell^{th}$ envelope.

\noindent\textbf{Step 2: Introduce a new stochastic process which bounds the total payment in a round}

Denote $\mathcal{F}_t$ as the filtration up to time $t$. Based on our algorithm, we know

\[ \Prob{\text{play a new arm}|\mathcal{F}_t} =
\begin{cases}
1       & \quad \text{if we incentivize}\\
\sum_{i\notin (V\cup S)} P(\AgentV{t} \cdot \ArmEV{t}{i}>\AgentV{t}\cdot \ArmEV{t}{j} \forall j\neq i)  & \quad \text{if }  S\setminus V\neq \emptyset 
\end{cases} \label{dom_stoc}
\] 
\hspace{1cm}$\geq (n-|V|)\times \frac{1}{n}=1-\frac{|V|}{n}$.

Let $(Z_{s,v,m}:m\geq 0)$ be a sequence of independent Bernoulli random variable with success probability $(1-\frac{v}{s})$. We will construct an alternative stochastic process for selecting which arm gets played that has the same distribution as the original process, but under which
\begin{align}
t_{s+1}-t_{s}\leq \bar{T}_{s}:=\sum_{v=0}^{\ARMNUM-1}\bar{T}_{s,v}, \nonumber 
\end{align}
where $\bar{T}_{s,v}:=\inf\{m\geq 0: Z_{s,v,m}=1\}+1$.

The new stochastic process will have the property that whenever $Z_{s,v(s,t),m(s,t)}=1$, we will play a new arm at time t for $t\in [t_{s-1}, t_{s}]$, where $v(s,t)$ is the number of unique arms played in round $s$ strictly before time $t$, and $m(s,t)$ is the number of times we have pulled a previously pulled arm for the current value of $v(s,t)$. At time $t$, to determine what arm to pull, calculate \eqref{dom_stoc} and let $q_t$ be the probability computed. Note $q_t\geq 1-\frac{v(s,t)}{s}$.

If $Z_{s,v(s,t),m(s,t)}$ is 1, decide to play a new arm. Otherwise, draw a second Bernoulli random variable with probability $\frac{q_t-(1-\frac{v(s,t)}{s})}{1-(1-\frac{v(s,t)}{s})}$, and if it is 1, decide to play a new arm, and otherwise decide to play an old arm. Note that

\begin{align}
\Prob{\text{play a new arm}|\mathcal{F}_{t}}=\left[1-\frac{v(s,t)}{s}\right]+\left[1-(1-\frac{v(s,t)}{s})\right]\times \frac{q_t-(1-\frac{v(s,t)}{s})}{1-(1-\frac{v(s,t)}{s})}=q_t.
\end{align}

Since we pull a new arm only when $Z_{s,v(s,t),m(s,t)}=1$, the number of times we pull an arm when $v(s,t)=v$ is bounded above by $\bar{T}_{s,v}$. Thus,

\begin{align}
t_{s+1}-t_{s}\leq \sum_{v=0}^{\ARMNUM-1}\bar{T}_{s,v}=\bar{T}_{s}.
\end{align}
Finally, we want to bound $\sum_{t}E[c(t)]$. Let $C(s)$ be the cost incurred in phase $s$. Thus $\sum_{t}E[c(t)]=\sum_{s}E[C(s)]$. We know
\begin{align}
C(s)&=\sum_{t=t_s}^{t_{s+1}-1}c(t)=\sum_{t=t_{s}}^{t_{s+1}-1}\sum_{\ell}\mathbbm{1}\{\omega\in L(\ell)\}c(t) \nonumber \\
&\leq \sum_{t=t_{s}}^{t_{s+1}-1}\sum_{l}\mathbbm{1}\{\omega\in L(\ell)\}\bar{c}(n,\ell) \nonumber \\
&\leq (t_{s+1}-t_{s})\sum_{\ell} \mathbbm{1}\{\omega\in L(\ell)\}\bar{c}(s,\ell) \nonumber \\
&\leq \bar{T}_{s}\sum_{l}\mathbbm{1}\{\omega\in L(\ell)\}\bar{c}(s,\ell). \nonumber
\end{align}

Since $\bar{T}_{s}$ is independent of $\omega$, we know
\begin{align}
E[C(s)]\leq E[\bar{T}_{s}]\sum_{\ell}\Prob{\omega\in L(\ell)}\bar{c}(s,\omega). \nonumber
\end{align}
\noindent\textbf{Step 3: Rewrite the total payment expression}.

Based on the above notations, we can rewrite the cumulative payment as follows:
\begin{align}
\sum_{t=1}^{\infty}c(t) =\sum_{s=1}^{\infty}C(s)
\leq  \sum_{s=1}^{\infty}\sum_{\ell=1}^{\infty}\bar{T}_{s}\mathbbm{1}\{\omega\in L(\ell)\}\bar{c}(s,\ell). \nonumber
\end{align}

Set $g(s,\ell)$ to be $\frac{12\sigma \ell}{s^{2/5}}$. Since if $|\Arm{i}{j}-\ArmE{t}{i}{j}|\leq \lambda$ is true $\forall i$, $\forall j$, then we know for those $\AgV\in \{\AgV:\AgV\cdot \ArmV{\Best{\AgV}}-\AgV \cdot \ArmV{\Second{\AgV}}> 2\Diam d\lambda\}$, they will correctly identify their best arm. Thus, if $|u_{i}^{j}-u_{i,t}^{j}|\leq \frac{12\sigma l}{s^{2/5}} \leq \frac{p^{-1}(\frac{\MinProb}{2})}{2\Diam d}$ $\forall i$ and $\forall j$, then the probability that an unincentivized agent would pull arm $i$ is at least $\frac{p}{2}$. Further, if time $t$ is in a phase $s$ that satisfies $s^{-1}\leq \MinProb/2$, then our algorithm will not incentivize pulling any arms. Denote $a_0=\frac{24\Diam d\sigma}{p^{-1}(\frac{\MinProb}{2})}$. In order to have $\frac{12\sigma l}{s^{2/5}}\leq \frac{p^{-1}(\frac{\MinProb}{2})}{2\Diam d}$, it is sufficient to have $s\geq \lceil (a_{0} l)^\frac{5}{2} \rceil$. In order to have $s^{-1}\leq \frac{\MinProb}{2}$, we need $s\geq \frac{2}{\MinProb}$. Denote $s_2=\frac{2}{\MinProb}$. Thus, we know we can only incur regret for sample paths $\omega$ in the $\ell^{th}$ envelope in the first $\max\{s_2,\lceil (a_0 \ell)^\frac{5}{2} \rceil\}$ rounds.

Thus,
\begin{align}
\sum_{t=1}^{\infty}c(t)\leq\sum_{\ell=1}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^\frac{5}{2} \rceil\}}\bar{c}(s,\ell)\mathbbm{1}\{\omega\in L[\ell]\}\bar{T}_{s}. \nonumber
\end{align}

Therefore,

\begin{align}
&E\left[\sum_{t=1}^{\infty}c(t)\right] \nonumber\\
\leq &\sum_{\ell=1}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^\frac{5}{2} \rceil\}}\bar{c}(s,\omega)\Prob{\omega \in L(\ell)}E[\bar{T}_{s}] \nonumber \\
=&\sum_{\ell=1}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^\frac{5}{2} \rceil\}}\left[\MAXR+2\Diam d\frac{12\sigma \ell}{s^{2/5}}\right]\Prob(\omega\in L[\ell])E[\bar{T}_{s}] \nonumber \\
\leq &\sum_{\ell=1}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^\frac{5}{2} \rceil\}}\left[\MAXR+24\Diam d\sigma l\right]\Prob(\omega\in L[\ell])E[\bar{T}_{s}]. \nonumber
\end{align}

\noindent\textbf{Step 4: Bound $\Prob{\omega\in L(\ell)}$}.

We now bound $\Prob{\omega\in L[\ell]}$ for $n\geq n_0$ and $\ell\geq 2$. As a reminder, we omit the dependency between $\epsilon_{i,t}^{j}$, $n$ and $\omega$. We know 
\begin{align}
&\Prob{\omega\in L[\ell]} \nonumber \\
=&\Prob{\omega\in L^{'}[\ell]}- \Prob{\omega\in L^{'}[\ell-1]} \nonumber \\
\leq & 1-\Prob{|\epsilon_{i,t}^{j}|<\frac{12\sigma (\ell-1)}{n^{2/5}}, \forall i,j} \nonumber \\
=&\Prob{\exists i,j, s.t. |\epsilon_{i,t}^{j}|\geq \frac{12\sigma (\ell-1)}{n^{2/5}}} \nonumber  \\
\leq &\sum_{i,j}\Prob{|\epsilon_{i,t}^{j}|\geq \frac{12\sigma (\ell-1)}{n^{2/5}}} \nonumber
\end{align}

Define $S_{i,t}^{j}=\frac{\NumPull{t}{i}\epsilon_{i,t}^{j}}{2\sigma}$, then we know $S_{i,t}^{j}$ is a summation of $1/2$ gaussian random numbers. Therefore,
\begin{align}
&\sum_{i,j}\Prob{|\epsilon_{i,t}^{j}|\geq \frac{12\sigma(\ell-1)}{n^{2/5}}} \nonumber \\ 
=&\sum_{i,j}\Prob{|S_{i,t}^{j}|\geq \frac{6\NumPull{t}{i}(\ell-1)}{n^{2/5}}} \nonumber \\
\leq &\sum_{i,j}\Prob{|S_{i,t}^{j}|\geq 6\NumPull{t}{i}^{3/5}(\ell-1)}. \nonumber
\end{align}

Based on Lemma~\ref{lemma:cal2}, we know
\begin{align}
&6\NumPull{t}{i}^{3/5}(\ell-1) \nonumber \\
=& 5\NumPull{t}{i}^{3/5} + \NumPull{t}{i}^{3/5}(\ell-1) \nonumber \\
\geq & \sqrt{0.6\NumPull{t}{i}\log(\log_{1.1}(\NumPull{t}{i})+1)} + \sqrt{(\ell-1)^2 \NumPull{t}{i}} \nonumber \\
\geq & \sqrt{0.6\NumPull{t}{i}\log(\log_{1.1}(\NumPull{t}{i})+1)+(\ell-1)^2 \NumPull{t}{i}}. \nonumber
\end{align}

Based on Lemma~\ref{lem:ACI-inequality}, we know
\begin{align}
& \Prob{\omega\in L(\ell)} \nonumber \\
\leq &\sum_{i,j}\Prob{|S_{i,t}^{j}|\geq 6\NumPull{t}{i}^{3/5}(\ell-1)} \nonumber \\
\leq & \sum_{i,j}24e^{-1.8(\ell-1)^2} = 24\ARMNUM d\exp(-1.8(\ell-1)^2). \nonumber
\end{align}

\noindent\textbf{Step 5: Final Step}

Thus,
\begin{align}
&\sum_{t=1}^{\infty}c(t) \nonumber \\
\leq&\sum_{\ell=1}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^{5/2} \rceil\}}\left[\MAXR+24\Diam d\sigma \ell\right]\Prob{\omega\in L[\ell])E[\bar{T}_{s}} \nonumber \\
\leq& \sum_{s=1}^{\max\{s_2,\lceil (a_0)^{5/2} \rceil\}}\left[\MAXR+24\Diam d\sigma \right]\Prob{\omega\in L[1]}E[\bar{T}_{s}] +\sum_{l=2}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^{5/2} \rceil\}}\left[\MAXR+24\Diam d\sigma \ell\right]\Prob{\omega\in L[\ell]}E[\bar{T}_{s}] \nonumber \\
\leq& \sum_{s=1}^{\max\{s_2,\lceil (a_0)^{5/2} \rceil\}}\left[R+24\Diam d\sigma \right]E[\bar{T}_{s}] +\sum_{\ell=2}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^{5/2} \rceil\}}\left[\MAXR+24\Diam d\sigma \ell\right]24\ARMNUM d e^{-1.8(\ell-1.9)^2}E[\bar{T}_{s}]\nonumber \\
\leq& \sum_{s=1}^{\max\{s_2,\lceil (a_0)^{5/2} \rceil\}}\left[\MAXR+24\Diam d\sigma \right]\ARMNUM s +\sum_{\ell=2}^{\infty}\sum_{s=1}^{\max\{s_2,\lceil (a_0 \ell)^{5/2} \rceil\}}\left[\MAXR+24\Diam d\sigma \ell\right]24 \ARMNUM d e^{-1.8(\ell-1.9)^2}\ARMNUM s \nonumber \\
\leq & \left[\MAXR+24\Diam d\sigma \right]\ARMNUM (\max\{s_2,\lceil (a_0)^{5/2} \rceil\})^2+\sum_{\ell=2}^{\infty}24\ARMNUM^2 d[\MAXR+24\Diam d\ell\sigma](\max\{s_2,\lceil (a_0 \ell)^{5/2} \rceil\})^2 e^{-1.8(\ell-1.9)^2} \nonumber \\
=& O(\frac{N}{\MinProb^5}+ \frac{\ARMNUM^2}{\MinProb^5}). \nonumber
\end{align}

\end{proof}

\section*{Calculation for Equation~\eqref{ex:limit}}

In this section, we prove $d(t)\leq \frac{z_{t,1}^{2}}{2}$ and $\lim_{t\rightarrow\infty}d(t)=\frac{z_{t,1}^2}{2}$. Denote $w(x)=\sqrt{x}$. Then we know $w(x+y)\leq \sqrt{x} + y\frac{1}{2\sqrt{x}}$ and $w(x+y)\geq \sqrt{x} + y\frac{1}{2\sqrt{x+y}}$ for $y\geq 0$. Thus, we know

\begin{align}
&\lim_{t\rightarrow\infty} t\left[1-\frac{\sqrt{t}+z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right] \nonumber \\
=& \lim_{t\rightarrow\infty} t\left[\frac{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}-\sqrt{t}-z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]  \nonumber \\
\geq & \lim_{t\rightarrow\infty} t\left[\frac{\sqrt{t}+z_{t,2}+z_{t,1}^{2}\frac{1}{2\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}-\sqrt{t}-z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]  \nonumber \\
= & \lim_{t\rightarrow\infty} t \times \frac{z_{t,1}^2}{2(z_{t,1}^2+(\sqrt{t}+z_{t,2})^2)} \rightarrow \frac{z_{t,1}^2}{2}. \nonumber 
\end{align}
and
\begin{align}
&\lim_{t\rightarrow\infty} t\left[1-\frac{\sqrt{t}+z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right] \nonumber \\
=& \lim_{t\rightarrow\infty} t\left[\frac{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}-\sqrt{t}-z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]  \nonumber \\
\leq & \lim_{t\rightarrow\infty} t\left[\frac{\sqrt{t}+z_{t,2}+z_{t,1}^{2}\frac{1}{2\sqrt{(\sqrt{t}+z_{t,2})^2}}-\sqrt{t}-z_{t,2}}{\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}}\right]  \nonumber \\
= & \lim_{t\rightarrow\infty} t \times \frac{z_{t,1}^2}{2\sqrt{z_{t,1}^2+(\sqrt{t}+z_{t,2})^2}(\sqrt{t}+z_{t,2})} \rightarrow \frac{z_{t,1}^2}{2}. \label{inequ:dct}
\end{align}   
Based on inequality~\eqref{inequ:dct}, we know $d(t)\leq \frac{z_{t,1}^{2}}{2}$ and $\lim_{t\rightarrow\infty}d(t)=\frac{z_{t,1}^2}{2}$.

\section*{Proof of Lemma~\ref{lem:n1-inequality}}

\begin{proof}
We first prove $2\sqrt{\frac{2}{1.8}s\log(T)}\geq \sqrt{0.6s\log(\log_{1.1}(s)+1)}$ for $s\geq 2$. This is true because
\begin{align}
&2\sqrt{\frac{2}{1.8}s\log(T)}\geq \sqrt{0.6s\log(\log_{1.1}(s)+1)} \nonumber \\
\Longleftarrow & 2\sqrt{s\log(T)}\geq \sqrt{s\log(\log_{1.1}(s)+1)} \nonumber \\
\Longleftarrow & 4\log(T)\geq \log(\log_{1.1}(s)+1) \nonumber \\
\Longleftarrow & T^4 \geq \log_{1.1}(s)+1 \nonumber \\
\Longleftarrow & s^4 \geq \log_{1.1}(s)+1, \nonumber
\end{align}
the derivative of $s^4-\log_{1.1}(s)-1$ is positive for $s\geq 2$ and $2^4-\log_{1.1}(s)-1>0$.

Therefore, we know
\begin{align}
&3\sqrt{\frac{2}{1.8}s\log(T)} \nonumber \\
= & 2\sqrt{\frac{2}{1.8}s\log(T)} + \sqrt{\frac{2}{1.8}s\log(T)} \nonumber \\
\geq & \sqrt{0.6s\log(\log_{1.1}(s)+1)}+\sqrt{\frac{2}{1.8}s\log(T)} \nonumber \\
\geq & \sqrt{0.6s\log(\log_{1.1}(s)+1)+\frac{2}{1.8}\log(T)s}. \nonumber 
\end{align}

\end{proof}