### 2018-02-16 technical-proofs.tex
cleaned up and tightened technical lemmas

\section{Proofs of Technical Lemmas}
\label{sec:n0-inequality-proof}
\label{sec:n1-inquality-proof}

In this section, we prove the two technical lemmas
\ref{lem:n0-inequality} and \ref{lem:n1-inequality}.
We restate both lemmas for convenience.

\dkcomment{I tightened these bounds a bit. Also, the proofs are quite
  a bit shorter now, so we might fold them into the main body?}

\begin{rtheorem}{Lemma}{\ref{lem:n0-inequality}}
For any $n \geq \LatePhase = \max (2, \frac{30 \sigma^3}{x^3})$, 
\begin{align*}
\sqrt{0.6 n \log (\log_{1.1}(n) + 1)} & \leq \frac{n x}{4 \sigma}. 
\end{align*}
\end{rtheorem}

\begin{proof}
By squaring the statement and rearranging, it is equivalent to show that

\begin{align*}
\frac{n}{\log(\log_{1.1}(n)+1)}
  & \geq \frac{9.6\sigma^2}{x^2}.
\end{align*}

A numerical calculation and derivative test shows that
$\frac{n}{\log(\log_{1.1}(n)+1)} \geq n^{2/3}$ for all $n \geq 2$,
and for $n \geq \frac{30 \sigma^3}{x^3}$, we get that
$n^{2/3} \geq \frac{9.6 \sigma^2}{x^2}$.
\end{proof}

\begin{rtheorem}{Lemma}{\ref{lem:n1-inequality}}
For any $T \geq 2$ and $1 \leq n \leq T$, we have
\begin{align*}
2\sqrt{\frac{2}{1.8} n \log(T)}
& \geq \sqrt{0.6 n \log(\log_{1.1}(n)+1) + \frac{2}{1.8} n \log(T)}.
\end{align*}
\dkcomment{This also holds for $n=1$, and I improved the constant.}
\end{rtheorem}

\begin{proof}
By squaring the desired inequality and canceling out common terms, the
statement to prove is equivalent to
$\frac{50}{9} \log T \geq \log(\log_{1.1}(n) + 1)$.
It is sufficient to show that
$\frac{50}{9} \log T \geq \log(\log_{1.1}(T) + 1)$,
which can be verified by numerical calculation for $T=2$ and a
derivative test.
\end{proof}