### 2018-06-06 algorithm_regret.tex
fixed proof of Theorem 3

\subsection{Bounding the Total Regret}
In bounding the total regret, because agents' types are from a compact
set by Assumption~\ref{A2}, the maximum regret in any one round is
bounded by a constant.
We use $\MAXR = \max_{\AgV, i,i'} \AgV \cdot (\ArmV{i} - \ArmV{i'})$
to denote this constant upper bound on the maximum regret that can be
incurred in any one time step.

\begin{emptyextraproof}{Theorem~\ref{rst:regret}}
Regret can arise in two ways:
(1) an agent was incentivized to pull a suboptimal arm, or
(2) an agent myopically pulled a suboptimal arm.
By Lemma~\ref{lem:numP}, Algorithm~\ref{alg:basic-incentivizing}
incentivizes agents at most 
$O\left( N\exp\left(\frac{2}{p}\right) \right)$
times in expectation, each time causing regret at most \MAXR,
for a total expected regret of at most
$O\left(\MAXR\cdot N\exp\left(\frac{2}{p}\right) \right)$.
For the rest of the proof, we focus on the regret incurred when agents
pull arms myopically and make mistakes.

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimates are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we define (with foresight) a phase-dependent
cutoff
$\EarlyCut{s} = \sqrt{\frac{128\log(\dkedit{s}) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}}$,
and consider a regret exceeding \EarlyCut{s} large.

For most of the proof, we will focus on the case $\EarlyCut{s} \leq \TieCutoff$.
The remaining phases are the ones with
$\dkedit{\frac{s}{\log s}} \leq \frac{128}{1.8} \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2$.
\dkedit{Because $\frac{s}{\log s} = \Omega(s^{2/3})$, there are at most
$O(\Diam^3 d^3 \sigma^3/\TieCutoff^3)$ such phases,}
each such phase lasts for at most $\ARMNUM \cdot \log(s)$ steps
in expectation by Lemma~\ref{lem:phase-length},
and each step incurs regret at most \MAXR.
Therefore, the total expected regret incurred in these phases is at most
\dkedit{$O(\ARMNUM \MAXR \cdot \Diam^3 d^3 \sigma^3/\TieCutoff^3
\cdot \log (\Diam^3 d^3 \sigma^3/\TieCutoff^3))
\leq O(\ARMNUM \MAXR \Diam^4 d^4 \sigma^4/\TieCutoff^4)
= O(\ARMNUM \MAXR)$,}
by crudely bounding \dkedit{$\log y \leq O(y^{1/3})$}.

For the remainder of the proof, we consider phases $s$
with $\EarlyCut{s} \leq \TieCutoff$.
We first consider agents \AgV incurring positive regret less than \EarlyCut{s}.
Let $A_s$ be the set of types satisfying 
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq \EarlyCut{s}$.
By Assumption~\ref{A1}, and because $\EarlyCut{s} \leq \TieCutoff$,
the total measure of $A_s$
is $\AlmostTied{\EarlyCut{s}} \leq \TieDensity \cdot \EarlyCut{s}$.
Then, by Lemma~\ref{lem:phase-length},
the expected number of pulls in phase $s$ by agents in $A_s$  
is bounded above by $\TieDensity \EarlyCut{s} \cdot \ARMNUM \log(s)$.
Any agent \AgV incurring positive regret less than $\EarlyCut{s}$
must have a type in $A_s$,
so the expected total regret from such agents,
summed over all phases, is at most
\begin{equation}
\sum_{s=1}^{T} \TieDensity \EarlyCut[2]{s} \cdot \ARMNUM \log(s).
\label{equ:small_regret_bound}
\end{equation}

We next bound the regret incurred in time steps with large regret.
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > t_{s+1}-t_{s}$.
By Lemma~\ref{lem:right-choice},
under \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}},
no agent at time $t$ will incur regret more than \EarlyCut{s}.

To bound the probability of \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}},
we use Lemma~\ref{lem:round-prob}
(with $x_s = x'_s = \frac{\EarlyCut{s}}{2 \Diam d}$
and a stopping time of $\tau_s^k$).
We first verify that this choice of $x_s, x'_s$
satisfies the assumption of Lemma~\ref{lem:round-prob} for all
$s \geq 2$,
i.e., that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x_n^2}{16 \sigma^2}}
\leq \frac{n x_n}{2 \sigma}$
for all $n \geq 2$.
Substituting that
$x_n = \frac{\EarlyCut{n}}{2 \Diam d} = \sqrt{\frac{32\dkedit{\log(n)} \cdot \sigma^2}{1.8 n}}$,
the left-hand side is

\[
  \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{2 \dkedit{\log n}}{1.8}}
  \; \leq \;
  \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{2 n \dkedit{\log n}}{1.8}},
\]
while the right-hand side is $\sqrt{\frac{8 n \dkedit{\log(n)}}{1.8}}$.
Squaring both sides and canceling out common terms,
the desired inequality is equivalent to
$\frac{50}{9} \dkedit{\log n} \geq \log(\log_{1.1}(n) + 1)$.
\dkdelete{Because $n \leq T$, it is sufficient to show that
$\frac{50}{9} \log T \geq \log(\log_{1.1}(T) + 1)$,}
This can be verified by numerical calculation for \dkedit{$n=2$} and a
derivative test.

Now, by Lemma~\ref{lem:round-prob} and a union bound over all arms $i$
and attributes $j$, we get that 
\begin{align*}
\Prob{\AccEU{\tau_s^k}{x_s}}
& \leq 24 \ARMNUM d \cdot \exp \left( \frac{-1.8 s \dkedit{x_s}^2}{16 \sigma^2} \right)
\; = \; \frac{24 \ARMNUM d}{\dkedit{s^2}}.
\end{align*}

In the low-probability event, we simply bound the maximum regret by \MAXR.
Summing over \dkedit{all} time periods,
the total expected regret from large-regret steps is at most
\dkedit{$24 \ARMNUM d \MAXR \cdot \frac{\pi^2}{6} = O(\ARMNUM \MAXR)$.}

Adding all four types of regret terms,
and substituting 
$\EarlyCut[2]{s} = \frac{128 \dkedit{\log(s)} \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most
$
O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
\dkedit{+ \ARMNUM \MAXR}
\dkedit{+ \ARMNUM \MAXR}
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} \EarlyCut[2]{s} \log(s) \right)\\
= 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \TieDensity \cdot \log^3(T) \right)$.\QED
\end{emptyextraproof}