### 2018-10-01 Discussion.tex
proposed general proof

\subsection{Tighter Bounds Under Additional Assumptions}
%: Discrete Preferences and Known \MinProb}

The proofs of Theorem~\ref{rst:budget} and Theorem~\ref{rst:regret}
incur exponential (in \ARMNUM) payment and regret in the initial
phases because the threshold $1/\log(s)$ required for incentivization
decreases slowly. 
This slow decrease is needed to bound the regret in the later phases
when the concentration inequality kicks in as in
Equation~(\ref{equ:small_regret_bound}).
\bccomment{In this secion, we first introduce a modified version of
\textit{payment-eligible} definition with general probability threshold, and analyze
its regret and payment bound. Then we discuss two restrictions under
which we can provide stronger bounds, avoiding this exponential dependence.}
%In this section, we discuss two restrictions under which we can
%modify the algorithm slightly and provide stronger bounds,
%avoiding this exponential dependence.
Both problem settings are special cases of the more general
setting previously considered.

\bccomment{
\subsubsection{General Algorithm}
In this section, we introduce a more general definition of payment-eligible.
More specifically, an arm $i$ is payment-eligible at time $t$ in phase $s$ when both of
the following are true:

\begin{itemize}
\item $i$ has been pulled at most $s$ times up to time $t$,
i.e., $\NumPull{t}{i} \leq s$.
\item Based on the current estimates \ArmEV{t}{i'} of all arms' attribute vectors,
    the probability of pulling arm $i$ is less than $\phi(s)$
\end{itemize}

The general algorithm operates the same as Algorithm~\ref{alg:basic-incentivizing},
except using the new definition of payment-eligible.

With this definition of payment-eligible arms and probability threshold $\phi(s)$, we can
prove

\begin{theorem}
\label{rst:general}
With the probability threshold $\phi(s)$, the expected total payment for 
Algorithm~\ref{alg:basic-incentivizing} is at most XXX, and the expected
cumulative regret up to time T is bounded above by XXX.
\end{theorem}
}

\subsubsection{Discrete Preferences}
\label{subsec:discrete}
Discrete preferences by agents are captured by the following
strengthening of Assumption~\ref{A1},
which states that the agents who are close to tied between
two arms have measure 0:

\begin{assumption}[Discrete Preferences]
\label{a:discrete}
There is a positive $\hat{z}$ such that
$\AlmostTied{\hat{z}} = 0$.
\end{assumption}

When Assumption~\ref{a:discrete} holds,
we restrict the payment-eligibility criterion by only incentivizing
arms with much smaller probability to be pulled:
\bccomment{we choose $\phi(s)=1/s$ (compared to $1/log(s)$ in Algorithm~\ref{alg:basic-incentivizing})}.

%an arm $i$ is payment-eligible at time $t$ in phase $s$ when both of
%the following are true:

%\begin{itemize}
%\item $i$ has been pulled at most $s$ times up to time $t$,
%i.e., $\NumPull{t}{i} \leq s$.
%\item Based on the current estimates \ArmEV{t}{i'} of all arms' attribute vectors,
%the probability of pulling arm $i$ is less than $1/s$
%(compared to $1/\log(s)$ in the general algorithm).
%\end{itemize}

We refer to this modified version of
Algorithm~\ref{alg:basic-incentivizing} as the
\emph{Discrete-Preference Algorithm.}
\bccomment{
    Plug $\phi(s)=1/s$ into Theorem~\ref{rst:general}, after some calculation (see details in XXX), we get:
%We outline a proof of the following result for this algorithm under
%Assumption~\ref{a:discrete} in Appendix~\ref{sec:discussion-proof1}.

\begin{corollary}
\label{rst:discrete}
Under Assumption~\ref{a:discrete}, the Discrete-Preference Algorithm
has expected payment bounded above by 
$O \left(\ARMNUM^2/\MinProb \right)$
and expected regret bounded above by $O(\ARMNUM / \MinProb)$.
\end{corollary}
}

%\begin{theorem}
%\label{rst:discrete}
%Under Assumption~\ref{a:discrete}, the Discrete-Preference Algorithm
%has expected payment bounded above by 
%$O \left(\ARMNUM^2/\MinProb \right)$
%and expected regret bounded above by $O(\ARMNUM / \MinProb)$.
%\end{theorem}

\subsubsection{Known \MinProb}
An alternative useful assumption is that \MinProb
(or a strictly positive lower bound on \MinProb) is known.

\begin{assumption}[Known $\MinProb$]
\label{a:known-p}
A strictly positive lower bound on \MinProb is known in advance.
\end{assumption}

When this assumption holds, we choose a different modification in the
definition of payment eligibility.
Let $s_0 = \exp(2/\MinProb)$ \bccomment{ and we set $\phi(s)=1/\log(s+s_0)$.}

%An arm $i$ is \emph{payment-eligible} at time $t$ (in phase $s$)
%if both:
%\begin{itemize}
%\item $i$ has been pulled at most
%$s$ times up to time $t$, i.e., $\NumPull{t}{i} \leq s$.
%\item Based on the current estimates \ArmEV{t}{i'} of all arms' attribute vectors,
%the probability of pulling arm $i$ is less than $1/\log(s+s_0)$.
%\end{itemize}

This knowledge of \MinProb allows the algorithm to incentivize
significantly fewer arm pulls.
We refer to this modified version of
Algorithm~\ref{alg:basic-incentivizing} as the \emph{Known-\MinProb
Algorithm.}

\bccomment{
Plug $\phi(s)=1/s$ into Theorem~\ref{rst:general}, after some calculation (see details in XXX), we get:
\begin{corollary}
\label{rst:known-p}
Under Assumption~\ref{a:known-p}, the Known-\MinProb Algorithm has
expected payment  bounded above by  
$O(\ARMNUM^2 \cdot \max(1,(\TieDensity/\MinProb)^{5/2}))$ 
and expected regret bounded above by
\begin{align*}
O \left( \frac{\ARMNUM^2}{\MinProb^2}
  +\frac{\ARMNUM \TieDensity \log^3(T)}{\MinProb}\right).
\end{align*}
\end{corollary}
}

%We outline a proof of the following result for this algorithm under
%Assumption~\ref{a:known-p} in Appendix~\ref{sec:discussion-proof2}.

%\begin{theorem}
%\label{rst:known-p}
%Under Assumption~\ref{a:known-p}, the Known-\MinProb Algorithm has
%expected payment  bounded above by  
%$O(\ARMNUM^2 \cdot \max(1,(\TieDensity/\MinProb)^{5/2}))$ 
%and expected regret bounded above by
%\begin{align*}
%$O \left( \frac{\ARMNUM^2}{\MinProb^2}
%  + \frac{\ARMNUM^2}{\hat{z}^2}+\frac{\ARMNUM}{\hat{z}^3}
%  +\frac{\ARMNUM \TieDensity \log^3(T)}{\MinProb}\right)$.
%  \end{align*}
%\end{theorem}